

\def\nameoflab{Temperature Trends in Sweden}
\def\coursecode{MNXB11}

\def\authorOne{Jens Hieronymus, Lukas Nord, Linn Preuss Jelvez}
\def\authorOneMail{}
\def\headerName{}

\def\supervisor{}
\def\supervisorMail{}

\documentclass[12pt,titlepage]{article}
\input{preamble.tex} 

% Use option lineno for line numbers 

\usepackage{float}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{graphicx}




%\title{Temperature Trends in Sweden}

%\author[1]{Jens Hieronymus}
%\author[1]{Lukas Nord}
%\author[1]{Linn Preuss Jelvez}
%\affil[1]{Group 8}

\begin{document}
\input{frontpage.tex}
%\maketitle
% \begin{abstract}
% Please provide an abstract of no more than 300 words. Your abstract should explain the main contributions of your article, and should not contain any material that is not included in the main text. 
% \end{abstract}



\flushbottom

\thispagestyle{empty}

\section*{ABSTRACT}
This project examines temperature trends in Sweden using open data from the Swedish Meteorological and Hydrological Institute covering the years 1850 to 2024. Three main analyses were carried out to study climate change, temperature patterns on specific days, and possible links between solar activity and temperature variation. The data were cleaned, processed, and analyzed using statistical methods and Fourier transforms.

The results show a clear warming trend across Sweden of about one degree Celsius per century. In Lund, the mean annual temperature has risen slightly, while the range between maximum and minimum temperatures has widened. This suggests that summers are becoming warmer and winters colder. The study of birthday temperatures showed a similar gradual increase over time with consistent patterns across both Lund and Luleå. The analysis of solar activity did not reveal any strong periodic trend related to the known eleven-year solar cycle, indicating that such effects are not easily detectable in Swedish temperature data.

Overall, the findings confirm a steady warming of the Swedish climate consistent with global patterns. The project demonstrates how open meteorological data and simple computational tools can be used to visualize and understand changes in climate.
%\keywords{Solar Cycle, Climate, Birthdays}

\input{Introduction}

\input{Data_Methods}

\input{Research_questions}

\newpage
\input{Results}
\input{climate}
\input{bdays}

\input{solar}


%\input{tex/Discussion}


%\bibliography{sample}

\end{document}
